---
title: "Class 14 Mini-project COVID-19 Vaccination Rates"
author: 'Xihan Zhou (PID: A15845684)'
date: '2022-03-03'
output: 
  pdf_document: 
    latex_engine: xelatex
---

# Getting Started
```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```


## Q1. What column details the total number of people fully vaccinated? 

persons_fully_vaccinated is the column that details the total number of people fully vaccinated.

## Q2. What column details the Zip code tabulation area? 

zip_code_tabulation_area is the column that details the Zip code tabulation area.

## Q3. What is the earliest date in this dataset? 
```{r}
vax$as_of_date[1]
```
2021-01-05 is the earliest date in this dataset.

## Q4. What is the latest date in this dataset?
```{r}
vax$as_of_date[nrow(vax)]
```
2021-03-01 is the latest date in this dataset.



```{r}
skimr::skim(vax)
```

## Q5. How many numeric columns are in this dataset? 

There are 9 numeric columns in this dataset.

## Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 
```{r}
sum( is.na(vax$persons_fully_vaccinated) )
```
There are 18338  “missing values” in the persons_fully_vaccinated column.

## Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 
```{r}
round(100*sum( is.na(vax$persons_fully_vaccinated) ) / length(vax$persons_fully_vaccinated), 2)
```
17.04% of the persons_fully_vaccinated values are missing.

## Q8. [Optional]: Why might this data be missing?
Some of the states might not report this kind of the data to the CDC so the data is missing.

# Working with dates
```{r}
library(lubridate)
```

```{r}
today()
```

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

## Q9. How many days have passed since the last update of the dataset? 
```{r}
(today() - vax$as_of_date[1]) - (vax$as_of_date[nrow(vax)] - vax$as_of_date[1])
```
2 days has passed since the last update of the dataset.

## Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 
```{r}
length(unique(vax$as_of_date))
```
There are 61 unique date in the dataset.

# Working with ZIP codes

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

# Focus on the San Diego area
```{r}
library(dplyr)
```

```{r}
# Subset to San Diego county only areas
sd <- vax[ vax$county == "San Diego" , ]
```
```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

## Q11. How many distinct zip codes are listed for San Diego County? 
```{r}
length(unique(sd$zip_code_tabulation_area))
```
There are 107 distinct zip codes listed for San Diego County.

## Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 
```{r}
sd[which.max(sd$age12_plus_population),]$zip_code_tabulation_area
```
92154 is the San Diego County Zip code area with the largest 12 + Population in this dataset.

## Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-03-01”? 
```{r}
sd.latest = filter(sd, as_of_date == "2022-03-01") 
mean(sd.latest$percent_of_population_fully_vaccinated, na.rm=T)
```
The overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-03-01” is 0.7053.


## Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-03-01”?
```{r}
hist(sd.latest$percent_of_population_fully_vaccinated, breaks = 30)
```

```{r}
library(ggplot2)

ggplot(sd.latest) +
  aes(percent_of_population_fully_vaccinated) + geom_histogram()
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

## Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:
```{r}
baseplot = ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x ="Date", y="Percent Vaccinated") +
  labs(title="Vaccination Rate for CA 92037 (UCSD)")
baseplot
```

## Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-03-01”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?
```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-03-01")

#head(vax.36)
```

```{r}
mean.36 = mean(vax.36$percent_of_population_fully_vaccinated, na.rm=T)
mean.36
```

Adding the lin3 showing the average vaccination rate for all zip code areas with a population just as large as 92037
```{r}
baseplot + geom_hline(yintercept = mean.36, linetype=2, color = "red")
```

## Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-03-01”?
```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```


## Q18. Using ggplot generate a histogram of this data.
```{r}
ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated) + geom_histogram() +
  labs(x="Percent of Population Fully Vaccinated in a ZIP code area", y="Count (ZIP code areas)") +
       labs(title="Histogram of Vaccination Rate Across San Diego County") +
  labs(subtitle="As of 2022-03-01")
```

## Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?
```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
The ZIP code 92109 is above the average value calculated above while 92040 is below the average value.

## Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.
```{r, warning=FALSE}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vacinated",
       title="Vaccination Rate Across CA",
       subtitle="Only areas with population above 36k are shown") +
  geom_hline(yintercept = mean.36, linetype=2, color = "red")
```

## Q21. How do you feel about traveling for Spring Break and meeting for in-person class afterwards?

Since on average area with 36k+ population have a percent vaccinated rate around 75, I feel safe traveling for Spring Break and meeting for in-person class afterward as long as we still keep the precautions for preventing COVID-19.
